---
title: "major_project_6_with_smote"
output: html_document
---
```{r}
# test branch

# clone not fork
```

```{r library}
library(tidyverse)
#library(here)      # directory referencing
#library(readxl)    # reading Excel files
#library(janitor)   # data cleaning 
#library(stringr)   # string manipulation
library(tidyr)     # new tidy functions
library(knitr) # kable
#library(modi) # ok for multivariate outlier detection
library(caret)# low variance filter
# missing values
#library(naniar)
#library(knitr)
#library(ggpubr) # ggplot arrangement
#ploting 
library(gridExtra)
library(kableExtra)
#outlier
#library(univOutl)
# tree methods
#library(tourr)
#library(RColorBrewer)
#library(plotly)
#library(htmltools)
library(performanceEstimation)# for SMOTE
library(rpart)
library(rpart.plot)
library(rattle) #fancyRpartPlot
library(Rtsne)
library(randomForest)
library(neuralnet)
library(e1071)# SVM regression
library(mltools)
library(data.table)
library(skimr)
library(smotefamily)
library(broom)
library(jtools)
```

```{r load data}
load("tech_data.Rdata") # load cleaned data from John's code, make sure you have the Rdata file within the working directory
```

```{r data selection by domain knowledge}
#dat<-tech_biom %>% filter (AGEC >= 19, AGEC<=64, SMKSTAT==5)  # filter age and smoke status
dat<-tech_biom %>% filter (AGEC >= 19, AGEC<=64)  # filter age only
var_list<-c("BMISC","SEX","AGEC","DIABBC","HCHOLBC","HSUGBC","HYPBC","PHDCMWBC","EXLWMBC","EXLWVBC", "SYSTOL","DIASTOL","TRIGRESB","CHOLRESB","LDLRESB","HBA1PREB","GLUCFREB","HDLCHREB","CVDMEDST","APOBRESB") # add/remove variables that are interested
dat2<-dat %>% select (var_list) # select columns that we are interested
dat2$EXLWMBC<-as.numeric(as.character(dat2$EXLWMBC)) # exerices time should be numeric 
dat2$EXLWVBC<-as.numeric(as.character(dat2$EXLWVBC)) # exerciese time should be numeric
str(dat2) # 7238 obs x 20 variables
```

```{r SMOTE function from DWmR package}
SMOTE <- function(form,data,
                  perc.over=200,k=5,
                  perc.under=200,
                  learner=NULL,...
                  )
  
  # INPUTS:
  # form a model formula
  # data the original training set (with the unbalanced distribution)
  # minCl  the minority class label
  # per.over/100 is the number of new cases (smoted cases) generated
  #              for each rare case. If perc.over < 100 a single case
  #              is generated uniquely for a randomly selected perc.over
  #              of the rare cases
  # k is the number of neighbours to consider as the pool from where
  #   the new examples are generated
  # perc.under/100 is the number of "normal" cases that are randomly
  #                selected for each smoted case
  # learner the learning system to use.
  # ...  any learning parameters to pass to learner
{

  # the column where the target variable is
  tgt <- which(names(data) == as.character(form[[2]]))
  minCl <- levels(data[,tgt])[which.min(table(data[,tgt]))]
  
  # get the cases of the minority class
  minExs <- which(data[,tgt] == minCl)

  # generate synthetic cases from these minExs
  if (tgt < ncol(data)) {
      cols <- 1:ncol(data)
      cols[c(tgt,ncol(data))] <- cols[c(ncol(data),tgt)]
      data <-  data[,cols]
  }
  newExs <- smote.exs(data[minExs,],ncol(data),perc.over,k)
  if (tgt < ncol(data)) {
      newExs <- newExs[,cols]
      data <- data[,cols]
  }
  
  # get the undersample of the "majority class" examples
  selMaj <- sample((1:NROW(data))[-minExs],
                   as.integer((perc.under/100)*nrow(newExs)),
                   replace=T)

  # the final data set (the undersample+the rare cases+the smoted exs)
  newdataset <- rbind(data[selMaj,],data[minExs,],newExs)

  # learn a model if required
  if (is.null(learner)) return(newdataset)
  else do.call(learner,list(form,newdataset,...))
}



# ===================================================
# Obtain a set of smoted examples for a set of rare cases.
# L. Torgo, Feb 2010
# ---------------------------------------------------
smote.exs <- function(data,tgt,N,k)
  # INPUTS:
  # data are the rare cases (the minority "class" cases)
  # tgt is the name of the target variable
  # N is the percentage of over-sampling to carry out;
  # and k is the number of nearest neighours to use for the generation
  # OUTPUTS:
  # The result of the function is a (N/100)*T set of generated
  # examples with rare values on the target
{
  nomatr <- c()
  T <- matrix(nrow=dim(data)[1],ncol=dim(data)[2]-1)
  for(col in seq.int(dim(T)[2]))
    if (class(data[,col]) %in% c('factor','character')) {
      T[,col] <- as.integer(data[,col])
      nomatr <- c(nomatr,col)
    } else T[,col] <- data[,col]
  
  if (N < 100) { # only a percentage of the T cases will be SMOTEd
    nT <- NROW(T)
    idx <- sample(1:nT,as.integer((N/100)*nT))
    T <- T[idx,]
    N <- 100
  }

  p <- dim(T)[2]
  nT <- dim(T)[1]

  ranges <- apply(T,2,max)-apply(T,2,min)
  
  nexs <-  as.integer(N/100) # this is the number of artificial exs generated
                                        # for each member of T
  new <- matrix(nrow=nexs*nT,ncol=p)    # the new cases

  for(i in 1:nT) {

    # the k NNs of case T[i,]
    xd <- scale(T,T[i,],ranges)
    for(a in nomatr) xd[,a] <- xd[,a]==0
    dd <- drop(xd^2 %*% rep(1, ncol(xd)))
    kNNs <- order(dd)[2:(k+1)]

    for(n in 1:nexs) {
      # select randomly one of the k NNs
      neig <- sample(1:k,1)

      ex <- vector(length=ncol(T))

      # the attribute values of the generated case
      difs <- T[kNNs[neig],]-T[i,]
      new[(i-1)*nexs+n,] <- T[i,]+runif(1)*difs
      for(a in nomatr)
        new[(i-1)*nexs+n,a] <- c(T[kNNs[neig],a],T[i,a])[1+round(runif(1),0)]

    }
  }
  newCases <- data.frame(new)
  for(a in nomatr)
    newCases[,a] <- factor(newCases[,a],levels=1:nlevels(data[,a]),labels=levels(data[,a]))

  newCases[,tgt] <- factor(rep(data[1,tgt],nrow(newCases)),levels=levels(data[,tgt]))
  colnames(newCases) <- colnames(data)
  newCases
}
```

```{r smote}
smoteData <- smote(HSUGBC ~ ., dat2, perc.over = 600,perc.under=1)
table(smoteData$HSUGBC)
```

```{r smote}
dat22<-dat2[,1:20] # do NOT include DIABBC and sores other than dia_score
dat22$HCHOLBC<-droplevels(dat22$HCHOLBC)
dat22<-na.omit(dat22)

table(dat22$HCHOLBC)
dat22<-na.omit(dat22)

dat22.balanced <-smote(HCHOLBC~., dat22, perc.over=500, perc.under=2)

table(dat22.balanced$HCHOLBC)


dat22.balanced<-dat22.balanced %>% mutate(
  cho_score= ifelse(HCHOLBC==5, 0, 1),
  
)

dat22.balanced$cho_score<-as.factor(dat22.balanced$cho_score)

# dat33.balanced<-dat33.balanced[,-4]
# dat33.balanced
# 
# model<-lm(dia_score~., data=dat33.balanced)

set.seed(2021)
training.samples <- dat22.balanced$cho_score %>% 
  createDataPartition(p = 0.8, list = FALSE)
train.data  <- dat22.balanced[training.samples, ]
test.data <- dat22.balanced[-training.samples, ]

dat22.balanced

#model<-nnet::multinom(cho_score~.-HCHOLBC, data=train.data)
model<-nnet::multinom(cho_score~.-HCHOLBC, data=train.data)
predict_class<-predict(model,newdata=test.data)


length(predict_class)
length(test.data$HCHOLBC)

summary(model)
data.frame(predict_class)

table(predict_class, test.data$cho_score)

mean(predict_class==test.data$cho_score)

training.samples <- dat22$HCHOLBC %>%
  createDataPartition(p = 0.8, list = FALSE)
train.data  <- dat22[training.samples, ]
test.data <- dat22[-training.samples, ]

model<-nnet::multinom(HCHOLBC~., data=train.data)

predict_class<-predict(model, newdata=test.data)
table(predict_class, test.data$HCHOLBC)

mean(predict_class==test.data$HCHOLBC)

table(dat22.balanced$cho_score)
# 
# summary(model)
```
# Dataset clean

## Encode Y:

### Y1 

Y1 is binary and contains only 2 classes: if never had disease, score = 0, else score = 1

```{r Y1}
Y1<-dat22.balanced %>% mutate(
  dia_score= as.factor(ifelse((DIABBC==5), 0, ifelse(DIABBC==3|DIABBC==2|DIABBC==1,1, NA))),
  cho_score= as.factor(ifelse((HCHOLBC==5), 0, ifelse(HCHOLBC==3|HCHOLBC==2|HCHOLBC==1,1, NA))),
  sug_score= as.factor(ifelse((HSUGBC==5),0, ifelse(HSUGBC==3|HSUGBC==2|HSUGBC==1,1,NA))),
  hyp_score= as.factor(ifelse((HYPBC==5), 0, ifelse(HYPBC==3|HYPBC==2|HYPBC==1,1, NA))),
  cvd_score= as.factor(ifelse((CVDMEDST==1|CVDMEDST==2|CVDMEDST==3),1,ifelse(CVDMEDST==4,0,NA))),
  final_score = dia_score+cho_score+sug_score+hyp_score+cvd_score
) %>% select(dia_score, cho_score, sug_score, hyp_score, cvd_score, final_score)

```

### Y2
Y2  contains 4 classes based on seriousness
```{r Y2}
Y2<-dat22.balanced %>% mutate(
  dia_score= ifelse(DIABBC==5, 0, ifelse(DIABBC==3, 1, ifelse(DIABBC==2, 2, ifelse(DIABBC==1, 3, NA)))),
  cho_score= ifelse(HCHOLBC==5, 0, ifelse(HCHOLBC==3, 1, ifelse(HCHOLBC==2, 2, ifelse(HCHOLBC==1, 3, NA)))),
  sug_score= ifelse(HSUGBC==5, 0, ifelse(HSUGBC==3, 1, ifelse(HSUGBC==2, 2, ifelse(HSUGBC==1, 3, NA)))),
  hyp_score= ifelse(HYPBC==5, 0, ifelse(HYPBC==3, 1, ifelse(HYPBC==2, 2, ifelse(HYPBC==1, 3, NA)))),
  cvd_score= ifelse((CVDMEDST==1|CVDMEDST==2|CVDMEDST==3),2,ifelse(CVDMEDST==4,1,NA)),
  final_score = dia_score+cho_score+sug_score+hyp_score+cvd_score
) %>% select(dia_score, cho_score, sug_score, hyp_score, cvd_score, final_score)

```

## Encode X:

### X2
```{r X2}
# Do not use magic number, 0,1,2,3 should be changed to meaningful levels later
X2<-dat22.balanced %>% mutate(
  Tri_score=as.factor(ifelse((TRIGRESB==1|TRIGRESB==2|TRIGRESB==3),0,ifelse((TRIGRESB==4|TRIGRESB==5),1,ifelse(TRIGRESB==97|TRIGRESB==98,NA,2)))),
  
  Chol_score=as.factor(ifelse((CHOLRESB==1|CHOLRESB==2|CHOLRESB==3),0,ifelse((CHOLRESB==4|CHOLRESB==5|CHOLRESB==6),1,ifelse(CHOLRESB==97|CHOLRESB==98,NA,2)))),
  
  LDL_score=as.factor(ifelse((LDLRESB==1|LDLRESB==2|LDLRESB==3|LDLRESB==4),0,ifelse((LDLRESB==5|LDLRESB==6|LDLRESB==7),1,ifelse(LDLRESB==97|LDLRESB==98,NA,2)))),
  
  Glu_score=as.factor(ifelse((GLUCFREB==4|GLUCFREB==5),0,ifelse((GLUCFREB==6|GLUCFREB==7),1,ifelse(GLUCFREB==97|GLUCFREB==98,NA,2)))),
  
  HDL_score=as.factor(ifelse((HDLCHREB==7|HDLCHREB==8),NA,ifelse((HDLCHREB==5|HDLCHREB==6),0,ifelse(HDLCHREB==1,2,1)))),
  
  
  ApoB_score=as.factor(ifelse((APOBRESB==1|APOBRESB==2|APOBRESB==3|APOBRESB==4),0,ifelse((APOBRESB==5),1,ifelse(APOBRESB==97|APOBRESB==98,NA,2)))),
  
  HbA1c_score=as.factor(ifelse((HBA1PREB==1|HBA1PREB==2),0,ifelse((HBA1PREB==3|HBA1PREB==4),1,ifelse(HBA1PREB==7|HBA1PREB==8,NA,2)))),

) %>% select (SYSTOL, DIASTOL,Tri_score, Chol_score, LDL_score, Glu_score, HDL_score, PHDCMWBC, EXLWMBC, EXLWVBC, ApoB_score, HbA1c_score)

str(X2)
```

### X1

Regroup ALL predictors into different levels such as "normal",  "high", based on domain knowledge. All predictors are categorical variables now.

```{r X1}
# Do not use magic number, 0,1,2,3 should be changed to meaningful levels later
X1<-dat22.balanced %>% mutate(
  Sys_score=as.factor(ifelse(SYSTOL<120,0,ifelse(SYSTOL<130,1,ifelse(SYSTOL<140,2,ifelse(SYSTOL>=998,NA,3))))),
  
  Dis_score=as.factor(ifelse(DIASTOL<80,0,ifelse(DIASTOL<90,1,ifelse(DIASTOL<100,2,ifelse(DIASTOL>=998,NA,3))))),
  
  Tri_score=as.factor(ifelse((TRIGRESB==1|TRIGRESB==2|TRIGRESB==3),0,ifelse((TRIGRESB==4|TRIGRESB==5),1,ifelse(TRIGRESB==97|TRIGRESB==98,NA,2)))),
  
  Chol_score=as.factor(ifelse((CHOLRESB==1|CHOLRESB==2|CHOLRESB==3),0,ifelse((CHOLRESB==4|CHOLRESB==5|CHOLRESB==6),1,ifelse(CHOLRESB==97|CHOLRESB==98,NA,2)))),
  
  LDL_score=as.factor(ifelse((LDLRESB==1|LDLRESB==2|LDLRESB==3|LDLRESB==4),0,ifelse((LDLRESB==5|LDLRESB==6|LDLRESB==7),1,ifelse(LDLRESB==97|LDLRESB==98,NA,2)))),
  
  Glu_score=as.factor(ifelse((GLUCFREB==4|GLUCFREB==5),0,ifelse((GLUCFREB==6|GLUCFREB==7),1,ifelse(GLUCFREB==97|GLUCFREB==98,NA,2)))),
  
  HDL_score=as.factor(ifelse((HDLCHREB==7|HDLCHREB==8),NA,ifelse((HDLCHREB==5|HDLCHREB==6),0,ifelse(HDLCHREB==1,2,1)))),
  
  Waist_score=as.factor(ifelse(SEX==1,ifelse(PHDCMWBC<102,0,ifelse(PHDCMWBC>=998,NA,1)),ifelse(PHDCMWBC<88,0,ifelse(PHDCMWBC>=998,NA,1)))),
  
  MBC_score=ifelse(EXLWMBC<150,0,1),
  
  VBC_score=ifelse(EXLWVBC<75,0,1),
  
  ApoB_score=as.factor(ifelse((APOBRESB==1|APOBRESB==2|APOBRESB==3|APOBRESB==4),0,ifelse((APOBRESB==5),1,ifelse(APOBRESB==97|APOBRESB==98,NA,2)))),
  
  HbA1c_score=as.factor(ifelse((HBA1PREB==1|HBA1PREB==2),0,ifelse((HBA1PREB==3|HBA1PREB==4),1,ifelse(HBA1PREB==7|HBA1PREB==8,NA,2)))),

) %>% select (Sys_score, Dis_score,Tri_score, Chol_score, LDL_score, Glu_score, HDL_score, Waist_score, MBC_score, VBC_score, ApoB_score, HbA1c_score)

X1$MBC_score<-as.factor(X1$MBC_score)
X1$VBC_score<-as.factor(X1$VBC_score)
str(X1)
```

###Smote data 2

# Data analysis
## Model2 Y1$sug_score~X1

```{r cho_score}
temp_dat<-data.frame(cbind(Y1$cho_score, X1))
temp_dat<-na.omit(temp_dat)
y<-temp_dat$Y1.cho_score
#x<-one_hot(as.data.table(temp_dat[, -1])) # first colume is y and the rest will be X
x<-temp_dat[,-1]
```
### Logistic regression Y1$sug_score~X1
```{r Y1$cho_score~X1 logsitic regression}
options(warn =-1) # do not show warning
# use one-hot encoding
#xx<-one_hot(as.data.table(temp_dat[, -1]))
xx<-x
# model selection (step/CV) + CV test results
repeats <- 10
res1 <- cv_penLogistic(y,xx,method="vanilla", repeats=repeats)
res2 <- cv_penLogistic(y,xx,method="fowardAIC", repeats=repeats)
res3 <- cv_penLogistic(y,xx,method="forwardBIC", repeats=repeats)
res4 <- cv_penLogistic(y,xx,method="stepwiseAIC", repeats=repeats)
res5 <- cv_penLogistic(y,xx,method="stepwiseBIC", repeats=repeats)
res6 <- cv_penLogistic(y,xx,method="ridge", repeats=repeats)
res7 <- cv_penLogistic(y,xx,method="lasso", repeats=repeats)
res8 <-cv_penLogistic(y,xx,method="firth",repeats=repeats)

save(res1,res2,res3,res4,res5,res6,res7,res8, file="logistic model2 Y1_cho~X1.Rdata")
```

```{r exhbit models and cv performance sug}
# compare error of above models

load("logistic model2 Y1_cho~X1.Rdata")

tab <- cbind(
  apply(res1[[1]],2,mean),
  apply(res2[[1]],2,mean),
  apply(res3[[1]],2,mean),
  apply(res4[[1]],2,mean),
  apply(res5[[1]],2,mean),
  apply(res6[[1]],2,mean),
  apply(res7[[1]],2,mean),
  apply(res8[[1]],2,mean))

colnames(tab) <- c("logistic",
                   "fwdAIC",
                   "fwdBIC",
                   "bwdAIC",
                   "bwdBIC",
                   "ridge",
                   "lasso",
                   "firth")

boxplot(tab)
datasummary_skim(data=data.frame(100*tab), fmt = "%.2f")

# extract coefficients of models
lcoef <- list(res1[[2]]$coef,
              res2[[2]]$coef,
              res3[[2]]$coef,
              res4[[2]]$coef,
              res5[[2]]$coef,
              drop(coef(res6[[2]], res6[[2]]$lambda.1se )),
              drop(coef(res7[[2]], res7[[2]]$lambda.1se )),
              res8[[2]]$coef
)

varnames <- unique(unlist(map(lcoef,names)))
tab = matrix(0, nrow = length(lcoef), ncol = length(varnames))
colnames(tab) = varnames
for (i in 1:length(lcoef)) 
  tab[i, names(lcoef[[i]])] = lcoef[[i]]

rownames(tab) <- paste("model",1:length(lcoef),sep="")
logistic_coef_table<-kable(t(tab))
logistic_coef_table

# model 6 ridge shows best result
drop(coef(res5[[2]], res6[[2]]$lambda.1se))

```
